---
title: "Convert csvs from Scanpy to Seurat object"
author: Natalie Charitakis
---

```{r}
library(Seurat)
library(stringr)
```

##Load in respective csvs


```{r}
obs <- read.csv('obs.csv')
var <- read.csv('var.csv')
X <- read.csv('X.csv', header=FALSE)
#delete first row of X dataframe
#colnames(X) <- NULL
X <- t(X)
#Add in numbers of names for the spot
colnames(X) <- 0:3029
```
##Generate metadata object to feed into Seurat object
```{r}
#get fake cell barcodes
cell_barcodes <- colnames(X)

#need get list of Orig.ident
orig.ident <- replicate(3030, "ff_breast_cancer")

#get number of counts per cell
nCount_RNA <- obs["n_counts"]

#get nFeature
nFeature_RNA <- obs["n_genes"]

#% MT
percent.mt <- obs["pct_counts_mt"]

#only extract clusters once as this will be populated twice in metadata
seurat_clusters <- obs["clusters"]

metadata <- data.frame(orig.ident, nCounts,nFeatures, percent.mt, seurat_clusters)
rownames(metadata) <- cell_barcodes
colnames(metadata) <- c("orig.ident", "nCount_RNA","nFeature_RNA","percent.mt","seurat_clusters")
metadata
```

#Now try creating Seurat object
```{r}
ff_breast_cancer_seurat <- CreateSeuratObject(X,
  project = "CreateSeuratObject",
  assay = "RNA",
  names.field = 1,
  meta.data = metadata,
)

#add active assay
ff_breast_cancer_seurat@active.assay <- c("RNA")
```

#get spatial coordinates
```{r}
#load in tissue position list
tissue_coordinates <- read.csv('V1_Breast_Cancer_Block_A_Section_1/spatial/tissue_positions_list.csv',header=FALSE)
colnames(tissue_coordinates) <- c("barcode","tissue","row","col","imagerow","imagecol")
tissue_coordinates

##sort tissue coordinates df by obs barcodes
sorted_barcodes <- obs$X

tissue_coordinates <- tissue_coordinates[match(sorted_barcodes, tissue_coordinates$barcode), ]
rownames(tissue_coordinates) <- tissue_coordinates$barcode
tissue_coordinates = subset(tissue_coordinates, select = -c(barcode)) 
tissue_coordinates

#create scale factors object to add
spot = 177.4817017129342
fiducial = 286.70121045935537
hires = 0.08250825
lowres = 0.024752475
scale_factors <- Seurat::scalefactors(spot=spot,fiducial = fiducial,hires = hires,lowres = lowres)

##add coordinates and name of assay
ff_breast_cancer_seurat@images$slice.1 = new(Class="VisiumV1",assay="RNA",key="image_",coordinates=tissue_coordinates,scale.factors=scale_factors)
ff_breast_cancer_seurat@images$slice.1@assay <- c("RNA")

```

##Load in same data from scanpy X file as scaled data
```{r}
ff_breast_cancer_seurat@assays$RNA@scale.data <- X 
```

##Extract HVGs from var file
```{r}
#select genes only where highly_variable is true
hvg_df <- subset(var, highly_variable == 'True')

#Now want to get this list of indices and add V infront of them
hvg_indices <- as.integer(rownames(hvg_df))
hvg_indices
```

```{r}

#replace each value in vector
for (i in hvg_indices){
  pos <- match(i,hvg_indices)
  print(hvg_indices[1])
  print(hvg_indices[pos])
  new_x <- gsub("^(.{0})(.*)$","\\V\\2",i)
  hvg_indices[pos] <- new_x
}

hvg_indices
```

##REplace variable features with one I've created
```{r}
ff_breast_cancer_seurat@assays$RNA@var.features <- hvg_indices

```

##Try finding spatially variable genes as is
```{r}
ff_breast_cancer_seurat <- FindSpatiallyVariableFeatures(ff_breast_cancer_seurat, assay = "RNA", features = VariableFeatures(ff_breast_cancer_seurat)[1:1000],
    selection.method = "markvariogram")
top.features <- head(SpatiallyVariableFeatures(ff_breast_cancer_seurat, selection.method = "markvariogram"), 1000)
saveRDS(top.features,"top_svgs_seurat_ff_breast_cancer.RDS")
```

##map genes back
#Get list of gene names
```{r}
library(stringr)

svgs_gene_names_indices <- c()

for (i in top.features){
  gene_name <- str_replace_all(i, "V", "")
  svgs_gene_names_indices <- append(svgs_gene_names_indices,gene_name)
}

svgs_gene_names_indices
```

#Extract the gene names from VAR by indices
```{r}
var_subset <- var[svgs_gene_names_indices, 1]
var_subset <- as.data.frame(var_subset)
write.csv(var_subset,'ff_breast_cancer_seurat_svgs_gene_names.csv')
```
